---
title: "CSC 587 Homework 1"
author: "Corey Osborne"
output:
  pdf_document: default
  html_notebook: default
---
We read Su_raw_matrix.txt into a data frame "su"
```{r}
su <- read.delim("Su_raw_matrix.txt")
```


Gets mean and standard deviation of Liver_2.CEL column,as well as the column mean and sum, and than print the results
```{r}
liver2_mean <- mean(su$Liver_2.CEL)
liver2_sd <- sd(su$Liver_2.CEL)

col_means <- colMeans(su)
col_sums <- colSums(su)

print(liver2_mean)
print(liver2_sd)
print(col_means)
print(col_sums)
```


Here we generate 10,000 random numbers to use for plotting histograms and normal distributions. We do one with a smaller standard deviation of 0.2, and another with a standard deviation of 0.5. 
```{r}
randomNum1 <- rnorm(10000, mean=0, sd=0.2)
hist(randomNum1, xlim=c(-5,5), main="mean=0, sd=0.2", xlab="Value", col="blue")

randomNum2 <- rnorm(10000, mean=0, sd=0.5)
hist(randomNum2, xlim=c(-5,5), main="mean=0, sd=0.5", xlab="value", col="green")

```


dat will be a sample dataset with two categories, A and B, which will contain 200 values each and then normally distributes, where A has a mean of 0 and B has a mean of 0.8.
```{r}
library(ggplot2)
dat <- data.frame(cond = factor(rep(c("A", "B"), each=200)), 
                  rating = c(rnorm(200),rnorm(200, mean=.8)))
```


This will produce an overlaid histogram, plotting 'rating' attribute, with bins of 0.5 from the dat sample dataset
```{r}
ggplot(dat, aes(x=rating, fill=cond)) +
  geom_histogram(binwidth=.5, alpha=.5, position="identity")
```


An interleaved histogram of the dat dataset, plotting 'rating' attribute (two side-by-side histograms)
```{r}
ggplot(dat, aes(x=rating, fill=cond)) + 
  geom_histogram(binwidth=.5, position="dodge")
```


Density plots of dat
```{r}
ggplot(dat, aes(x=rating, colour=cond)) + geom_density()
```


Density plots of dat with fill
```{r}
ggplot(dat, aes(x=rating, fill=cond)) + geom_density(alpha=.3)
```


Instead of using random sample set (dat), we'll use a diabetes dataset produced from diabetes_train.csv
```{r}
diabetes <- read.csv("diabetes_train.csv")
```


Overlaid histogram of diabetes dataset, plotting mass attribute
```{r}
ggplot(diabetes, aes(x=mass, fill=class)) +
  geom_histogram(binwidth=.5, alpha=.5, position="identity")
```


Interleaved histogram of diabetes dataset, plotting mass attribute
```{r}
ggplot(diabetes, aes(x=mass, fill=class)) + 
  geom_histogram(binwidth=.5, position="dodge")
```


Diabetes dataset density plot
```{r}
ggplot(diabetes, aes(x=mass, colour=class)) + geom_density()
```


Diabetes dataset density plot with fill
```{r}
ggplot(diabetes, aes(x=mass, fill=class)) + geom_density(alpha=.3)
```

```{r}
library(tidyverse)
passengers <- read.csv("titanic.csv")
```


Here we remove any rows with missing values and provide a summary of statistical information
```{r}
passengers %>% drop_na() %>% summary()
```


This filters and then creates a new dataset containing only male passengers (only 10 displayed)
```{r}
passengers %>% filter(Sex == "male") %>% head(10)
```


We then sort all passengers by Fare, in descending order (the highest-paying passengers will appear first) ) (only 10 displayed)
```{r}
passengers %>% arrange(desc(Fare)) %>% head(10)
```


This calculates the total number of family members onboard, where Parch is parents/children, and SibSp is siblings/spouse (only 10 displayed)
```{r}
passengers %>% mutate(Famsize = Parch + SibSp) %>% head(10)
```


Here we check the average ticket fare by gender, as well as find the total number of survivors per gender
```{r}
passengers %>% group_by(Sex) %>% 
  summarise(meanFare = mean(Fare), numSurv = sum(Survived))
```


Here we use quantile to calculate the 10th, 30th, 50th, and 60th percentiles of the skin attribute from the diabetes dataset
```{r}
percentiles <- quantile(diabetes$skin, probs = c(0.1, 0.3, 0.5, 0.6))
print(percentiles)
```
